General constraints on magic distillation protocols from statistical mechanics

### 2021-05-04 appendix.tex

\section{Properties of Wigner distributions}
\label{app:wigner}

Here we present important properties of the phase-point operators and the Wigner distribution that are used throughout the paper.

\begin{proposition}\label{thm:aproperties}
    For any dimension $d$, the phase-point operators satisfy:
    \begin{enumerate}
        \item[(i)]\label{en:a1} Hermiticity and unitarity: $A_{\bmx}^\dagger = A_{\bmx} = A_{\bmx}^{-1}$;
	    \item[(ii)]\label{en:a2} Closure under transposition: $A_{(x, p)}^T = A_{(x, p)}$;
	    \item[(iii)]\label{en:a3} Unit trace for odd $d$: $\tr[A_{\bmx}] = 1$;
	    \item[(iv)]\label{en:a4} Completeness relation: $\sum_{\bmz \in \cal{P}_d} A_{\bmz} = d\id$;
	    \item[(i)]\label{en:a5} Orthogonality: $\tr[A_{\bmx}^\dagger A_{\bm{x'}}] = d \delta_{\bmx,\bm{x'}}$.
	\end{enumerate}
\end{proposition}
\begin{proof}
	All properties follow from the definition in~\cref{eq:ax} along with properties of the displacement operator $D_{\bmx}$ and can be found in the literature, e.g.~\cite{cit:veitch,Vourdas_2004,cit:gross3}
\end{proof}

@@

\subsection{Comments on monotones and fragements}
The development of any monotone $\M(\rho)$ for a magic theory $\R$ can be viewed abstractly as
\begin{equation}
\M : \mbox{(pre-order of states)} \mapsto \mbox{(total order of numbers).} \nonumber
\end{equation}
More precisely, within the theory we order magic states as $\rho_1 \succ \rho_2$ whenever there is an operation $\E \in \O$ such that $\E(\rho_1) = \rho_2$. In general many states will not be interconvertible in both directions, so this ordering is a pre-order in the sense that incomparable states exist. A magic monotone projects this pre-order of states onto the real number line, and, through this lens, any two states are directly comparable in terms of their mapped values $\M(\rho_1)$, $\M(\rho_2)$. Of course, the price is that the monotone provides only a necessary, but not sufficient, condition for the transformation $\rho_1 \rightarrow \rho_2$.

The restriction of a magic theory $\R$ to some fragment $\R_\sigma$ can now be interpreted as a generalisation of the magic monotone construction: instead of mapping the pre-order of states in $\R$ onto the number line, we instead map this pre-order onto \emph{a simpler pre-order of states}. Therefore, we are retaining more of the original parent theory, while working in a scenario that admits greater structure and simplicity. We can view this as a mapping $\P_\sigma: \R \rightarrow \R_\sigma$ that restricts to a sub-theory (a $\sigma$--fragment) of the parent theory $\R$. Thus,
\begin{equation}
\P_\sigma: \mbox{(pre-order)} \longrightarrow \mbox{(simpler pre-order).}
\end{equation}
By framing the projection onto a fragment in this light, we can view Theorem \ref{thm:frag} as a statement that the collection $\{\P_\sigma\}_{\sigma \in \F}$ provides a complete set of generalized monotones for the theory.
\subsection{Simple--majorization equivalence conditions}

In the unital fragment, namely the limit of infinite temperature, $\beta = 0$, the free state is the maximally mixed state $\frac{1}{n}\id$ with uniform Wigner distribution $\frac{1}{n}\bm{1} = (\frac{1}{n},\dots,\frac{1}{n})$.
This fragment is governed by simple majorization and we first prove strong equivalences for this type of majorization.
\begin{proposition}\label{prop:major}
Given $\bmx, \bmy \in \mathbb{R}^n$ the following statements are equivalent:
 \begin{enumerate}
	\item[(i)]\label{en:m1} $\bmx \prec \bmy$;
	\item[(ii)]\label{en:m2} $\bmx = B \bmy$ for bistochastic $B$;
	\item[(iii)]\label{en:m3} $\sum\limits_{i=1}^n \abs{x_i - t} \leq \sum\limits_{i=1}^n \abs{y_i - t}$ for all $t \in \mathbb{R}$;
	\item[(iv)]\label{en:m4} $\sum\limits_{i=1}^n (x_i - t)^+ \leq \sum\limits_{i=1}^n (y_i - t)^+$ for all $t \in \mathbb{R}$ and \vspace{5pt}\\ $\sum\limits_{i=1}^n x_i = \sum\limits_{i=1}^n y_i$, where $(x)^+ = \max{\{x, 0\}}$;
	\item[(v)]\label{en:m5} $L_{\bmx}(t) \leq L_{\bmy}(t)$ for $t\in [0,1)$	and \vspace{5pt}\\ $L_{\bmx}(1) = L_{\bmy}(1)$, where $L_{\bmz}(t) = L_{\bmz|\frac{1}{n}\bm{1}}(t)$.
 \end{enumerate}
\end{proposition}
	The equivalence between (i) and (ii) is the statement of the Hardy, Littlewood and Polya theorem and is taken as our definition of majorisation in the main text.

	The rest of the properties are well-known and proofs can be found in~\cite{cit:marshall,cit:bhatia,cit:nielsen,cit:lostaglio}.

\begin{lemma} Let $\p$ be a quasi-probability distribution and let $\r$ be a probability distribution with strictly non-zero components. Let $a > 0$ and $b \in \mathbb{R}$ then $L_{a\p + b \r | \r} (x) = a L_{\p |\r}(x) + b x$.
\end{lemma}\label{lemma:Lorenz_linearity}
\begin{proof} The Lorenz curve of $a\p + b \r$ relative to $\r$ passes through $(0,0)$ and the points $(\sum_{i=1}^k{r_{\pi(i)}}, \sum_{i=1}^k(a \p + b \r)_{\pi(i)})$ where $\pi$ is the permutation that puts $(a p_i/r_i + b)$ in non-increasing order. Since $a > 0$, the permutation $\pi$ puts  $(p_i/r_i)$ in non-increasing order too. We thus have
\begin{align*}
&\left( \sum_{i=1}^kr_{\pi(i)}, \sum_{i=1}^k(a \p + b \r)_{\pi(i)} \right) = \\ 
&\left( \sum_{i=1}^k r_{\pi(i)},a \sum_{i=1}^k  p_{\pi(i)} + b\sum_{i=1}^k r_{\pi(i)} \right) \nonumber,
\end{align*}
therefore the value of the Lorenz function at each potential elbow point $x_k = \sum_{i=1} ^kr_{\pi(i)}$ is given by
\begin{align}
&L_{a \p +b \r|\r} (x_k) = a L_{\p|\r} (x_k) + b L_{\r|\r}(x_k) = \nonumber\\
&a L_{\p|\r} (x_k) + b x_k,
\end{align}
so we have $L_{a\p  + b\r|\r} (x) = a L_{\p |\r}(x) + b x$ for any $x \in [0,1]$ due to linearity.
\end{proof}

@@

\subsection{Embedding map}

Any $\r$--majorization problem can be rephrased as a simple majorization problem in a higher dimensional space via the embedding map.
\begin{definition}
  The embedding map $\Gamma_{\r}:\mathbb{R}^n \mapsto \mathbb{R}^N, N = \sum\limits_{i=1}^n r_i$ is the function
  \begin{equation}
    \Gamma_{\r}(\bmw) = \bigoplus_{i=1}^n w_i\  \frac{1}{r_i}\bm{1},
  \end{equation}
where $\bm{1}/r_i$ is the $r_i$--dimensional uniform distribution.
The left inverse $\Gamma_{\r}^{-1}: \mathbb{R}^N \mapsto \mathbb{R}^n$ is defined to sum up the elements in each block of $\Gamma_{\r}(\bmw)$, so that
  \begin{equation}
     \Gamma_{\r}^{-1}(\oplus_{i=1}^n w_i \bm{1}/r_i) = \bmw.
  \end{equation}
  This is not a right inverse, because $\Gamma_{\r}$ is not surjective.
\end{definition}
The direct sum simply amounts to listing the uniform distributions one after the other.
The embedding map maps the Gibbs distribution to the uniform distribution, $\Gamma_{\r}(\r) = \bm{1}/N$.
Then, a non-increasing ordering $\Gamma_{\r}(\bmz)^\downarrow$ in the new space, corresponds to the so-called ``$\beta$-ordering'' of the original vector denoted by the permutation $\pi$ in~\cref{eq:lc}, mapping $(w_i/r_i) \mapsto (w_i/r_i)^\downarrow$ for all $i=1,\dots,n$.

\subsection{\textit{r}-majorization equivalence conditions}

We take the opportunity here	 to simply list useful equivalent statements for $\r$--majorisation.

\begin{proposition}\label{prop:rmajor}
Given $\bmx, \bmy, \r \in \mathbb{R}^n$, such that the components of $\r$ are positive, the following statements are equivalent:
  \begin{enumerate}
    \item[(i)] $\bmx \prec_{\r} \bmy$;
    \item[(ii)] $\Gamma_{\r}({\bmx}) \prec \Gamma_{\r}({\bmy})$;
    \item[(iii)]\label{en:tm3} $\sum\limits_{i=1}^n \abs{x_i - r_i t} \leq \sum\limits_{i=1}^n \abs{y_i - r_i t}$ for all $t \in \mathbb{R}$;
    \item[(iv)] $\sum\limits_{i=1}^n (x_i - r_i t)^+ \leq \sum\limits_{i=1}^n (y_i - r_i t)^+$ for all $t \in \mathbb{R}$ and \vspace{5pt}\\ $\sum\limits_{i=1}^n x_i = \sum\limits_{i=1}^n y_i$;
    \item[(v)] $L_{\bmx|\r}(t) \leq L_{\bmy|\r}(t)$ for $t\in [0,1)$ and \vspace{5pt}\\ $L_{\bmx|\r}(1) = L_{\bmy|\r}(1)$.
  \end{enumerate}
\end{proposition}
\begin{proof}$ $\vspace{-12pt}\\
\begin{enumerate}
	\item[1$\leftrightarrow2$]
	Suppose there exists a $\r$--stochastic $A$ such that $\bmx = A\bmy$ and let $B = \Gamma_{\r} \circ A \circ \Gamma_{\r}^{-1}$.
$B$ is a $N$--dimensional bistochastic matrix, since composition of stochastic matrices is stochastic and $(\Gamma_{\r} \circ A \circ \Gamma_{\r}^{-1})\big(\frac{1}{N}\bm{1}\big) = (\Gamma_{\r} \circ A) (\r) = \Gamma_{\r}(\r) = \frac{1}{N}\bm{1}$. Then, $B$ maps $\Gamma_{\r}({\bmy})$ to $\Gamma_{\r}({\bmx})$.

	Conversely, given $B$, let $A = \Gamma_{\r}^{-1} \circ B \circ \Gamma_{\r}$.
	Similarly, $A$ is a $\r$--stochastic matrix that maps $\bmy$ to $\bmx$.
	\item[$2\leftrightarrow3$] These \hspace{1pt} three \hspace{1pt} statements \hspace{1pt} are \hspace{1pt} equivalent \hspace{1pt} to \hspace{1pt} state-
	\item[$2\leftrightarrow4$] \vspace{-7pt} ments (iii), (iv), (v) \hspace{1pt} in~\cref{prop:major} respectively, 
	\item[$2\leftrightarrow5$] \vspace{-7pt} for the embedded vectors $\Gamma_{\r}({\bmx}), \Gamma_{\r}({\bmy})$, which becomes apparent by rewriting
	\begin{align*}
	&\hspace{15pt} \sum\limits_{i=1}^n \abs{x_i - r_i t} = \sum\limits_{i=1}^n r_i \abs{\frac{x_i}{r_i} - t} = \sum\limits_{i=1}^N \abs{\Gamma_{\r}(\bmx)_i - t}, \\[3pt]
	&\hspace{15pt} \sum\limits_{i=1}^n (x_i - r_i t)^+ = \sum\limits_{i=1}^N (\Gamma_{\r}(\bmx)_i - t)^+ \text{ and} \\[6pt]
	&\hspace{15pt} L_{\bmx|\r}(t) = L_{\Gamma_{\r}(\bmx)}(t),
	\end{align*} 
and similarly for the right hand sides of the inequalities.
\end{enumerate}
\vspace{-19pt}
\end{proof}

@@

\subsection{Lorenz curve coordinates in the unital fragment}\label{app:lcsu_coord}
The Wigner distribution of the $n$--copy qutrit maximally mixed state $\left(\id/3\right)^{\otimes n}$ is the uniform probability distribution over the phase space, consisting of $9^n$ components equal to $9^{-n}$.
The Wigner distribution of the 1-copy $\epsilon$--noisy Strange state $\rho_{\rm{S}}(\epsilon)$ in the unital fragment consists of some permutation of a single negative component
\begin{equation}
	- v(\epsilon) \coloneqq - \left( \frac{1}{3} -\frac{4}{9}\epsilon \right),
\end{equation} 
and $8$ positive components
\begin{equation}
	u(\epsilon) \coloneqq \frac{1}{6} -\frac{1}{18}\epsilon.
\end{equation}
where in the unital fragment we need the condition $0 \leq \epsilon < 3/4$, so that the state contains some Wigner negativity ($-v < 0$).
It is also clear that $v \geq u$ in the interval $0 \leq \epsilon \leq 3/7$, while $u > v$ in the interval $3/7 < \epsilon < 3/4$.

@@

\begin{theorem}
	Consider a magic distillation protocol transforming noisy Strange state
	\begin{equation*}
		\rho_S(\epsilon)^{\otimes n} \longrightarrow \E(\rho_S(\epsilon)^{\otimes n})=\rho_S(\epsilon')^{\otimes m} 
	\end{equation*}
	with $n, m \gg 1$.

>>>>

\newpage
\section{Properties of Wigner distributions}
\label{app:wigner}

Here we present basic properties of the phase-point operators and the Wigner distribution that are used throughout the paper.

\begin{proposition}\label{thm:aproperties}
    For any dimension $d$, the phase-point operators satisfy:
    \begin{enumerate}
        \item[(i)]\label{en:a1} Hermiticity and unitarity: $A_{\bmx}^\dagger = A_{\bmx} = A_{\bmx}^{-1}$;
	    \item[(ii)]\label{en:a2} Closure under transposition: $A_{(x, p)}^T = A_{(x, p)}$; \ddd{[FIX]}
	    \item[(iii)]\label{en:a3} Unit trace for odd $d$: $\tr[A_{\bmx}] = 1$;
	    \item[(iv)]\label{en:a4} Completeness relation: $\sum_{\bmz \in \cal{P}_d} A_{\bmz} = d\id$;
	    \item[(i)]\label{en:a5} Orthogonality: $\tr[A_{\bmx}^\dagger A_{\bm{x'}}] = d \delta_{\bmx,\bm{x'}}$.
	\end{enumerate}
\end{proposition}
\begin{proof}
	All properties follow from the definition in~\cref{eq:ax} along with properties of the displacement operator $D_{\bmx}$ and can be found in the literature, e.g.~\cite{cit:veitch,Vourdas_2004,cit:gross3}
\end{proof}

@@

We now give the following equivalent formulations of $d$--majorization.

\begin{proposition}\label{prop:rmajor}
Given $\bmx, \bmy, \r \in \mathbb{R}^n$, such that the components of $\r$ are positive, the following statements are equivalent:
  \begin{enumerate}
    \item[(i)] $\bmx \prec_{\r} \bmy$;
    \item[(ii)] $\Gamma_{\r}({\bmx}) \prec \Gamma_{\r}({\bmy})$;
    \item[(iii)]\label{en:tm3} $\sum\limits_{i=1}^n \abs{x_i - r_i t} \leq \sum\limits_{i=1}^n \abs{y_i - r_i t}$ for all $t \in \mathbb{R}$;
    \item[(iv)] $\sum\limits_{i=1}^n (x_i - r_i t)^+ \leq \sum\limits_{i=1}^n (y_i - r_i t)^+$ for all $t \in \mathbb{R}$ and \vspace{5pt}\\ $\sum\limits_{i=1}^n x_i = \sum\limits_{i=1}^n y_i$;
    \item[(v)] $L_{\bmx|\r}(t) \leq L_{\bmy|\r}(t)$ for $t\in [0,1)$ and \vspace{5pt}\\ $L_{\bmx|\r}(1) = L_{\bmy|\r}(1)$.
  \end{enumerate}
\end{proposition}
The proofs for these can be found in~\cite{cit:marshall,cit:bhatia,cit:nielsen,cit:lostaglio} and references therein.

The following lemma is used in the text to relate relative majorization of quasi-distributions to their Lorenz curves.
\begin{lemma} Let $\p$ be a quasi-probability distribution and let $\r$ be a probability distribution with strictly non-zero components. Let $a > 0$ and $b \in \mathbb{R}$ then $L_{a\p + b \r | \r} (x) = a L_{\p |\r}(x) + b x$.
\end{lemma}\label{lemma:Lorenz_linearity}
\begin{proof} The Lorenz curve of $a\p + b \r$ relative to $\r$ passes through $(0,0)$ and the points $(\sum_{i=1}^k{r_{\pi(i)}}, \sum_{i=1}^k(a \p + b \r)_{\pi(i)})$ where $\pi$ is the permutation that puts $(a p_i/r_i + b)$ in non-increasing order. Since $a > 0$, the permutation $\pi$ puts  $(p_i/r_i)$ in non-increasing order too. We thus have
\begin{align*}
&\left( \sum_{i=1}^kr_{\pi(i)}, \sum_{i=1}^k(a \p + b \r)_{\pi(i)} \right) = \\ 
&\left( \sum_{i=1}^k r_{\pi(i)},a \sum_{i=1}^k  p_{\pi(i)} + b\sum_{i=1}^k r_{\pi(i)} \right) \nonumber,
\end{align*}
therefore the value of the Lorenz function at each potential elbow point $x_k = \sum_{i=1} ^kr_{\pi(i)}$ is given by
\begin{align}
&L_{a \p +b \r|\r} (x_k) = a L_{\p|\r} (x_k) + b L_{\r|\r}(x_k) = \nonumber\\
&a L_{\p|\r} (x_k) + b x_k,
\end{align}
so we have $L_{a\p  + b\r|\r} (x) = a L_{\p |\r}(x) + b x$ for any $x \in [0,1]$ due to linearity.
\end{proof}

@@

@@

\ddd{[Remainder bounds via integrals? This might be of use for others, if they wanted to build on this line.]}
\subsection{Lorenz curve coordinates in the unital fragment}\label{app:lcsu_coord}
The Wigner distribution of the $n$--copy qutrit maximally mixed state $\left(\id/3\right)^{\otimes n}$ is the uniform probability distribution over the phase space, consisting of $9^n$ components equal to $9^{-n}$.
The Wigner distribution of the 1-copy $\epsilon$--noisy Strange state $\rho_{\rm{S}}(\epsilon)$ in the unital fragment consists of some permutation of a single negative component
\begin{equation}
	- v(\epsilon) \coloneqq - \left( \frac{1}{3} -\frac{4}{9}\epsilon \right),
\end{equation} 
and $8$ positive components
\begin{equation}
	u(\epsilon) \coloneqq \frac{1}{6} -\frac{1}{18}\epsilon.
\end{equation}
where in the unital fragment we need the condition $0 \leq \epsilon < 3/4$, so that the state contains some Wigner negativity ($-v < 0$).
It is also clear that $v \geq u$ in the interval $0 \leq \epsilon \leq 3/7$, while $u > v$ in the interval $3/7 < \epsilon < 3/4$.

@@

\ddd{[Do we need to keep any of the details in this section? If not then delete it entirely.]}
\begin{theorem}
	Consider a magic distillation protocol transforming noisy Strange state
	\begin{equation*}
		\rho_S(\epsilon)^{\otimes n} \longrightarrow \E(\rho_S(\epsilon)^{\otimes n})=\rho_S(\epsilon')^{\otimes m} 
	\end{equation*}
	with $n, m \gg 1$.

### 2021-05-04 magic_paper.tex

The above analysis makes simplifying assumptions that could easily be dropped, at the price of more complex expressions. For example, we can obtain general expressions without any assumptions on $\epsilon,\epsilon'$ or drop the Clifford change in basis. 
We could also perform similar analysis for general qudits, and different choices of magic states. It might also be of interest consider other choices of reference states that may include more appropriate hardware physics, for example for photonic set-ups.

The only non-trivial assumption is the assumption that we can ignore correlations in the thermodynamic limit. This is a common situation in thermodynamics, where correlations are often sub-linear in the particle number in the macroscopic limit. One could perturb about this scenario, and make use of variational tools such as the Bogoliubov inequality \ddd{[cite]} for approximating the free energy of a system, to obtain bounds in a similar form.
\begin{figure}[t!]
    \centering
    \includegraphics[scale=0.4]{figs/rate_scatter.pdf}
    \caption{\textbf{Temperature-dependent bounds for magic distillation.}
  Shown is a contour plot of  $R(\epsilon, \beta)$ for $H=H'$, where $\beta$ is an inverse temperature and $\epsilon$ is the depolarising noise of the magic state. The relevant data from the spectrum of the qutrit operator $H$ enters in via $\epsilon_\star(\beta)$ and $\beta_\star$.  The vertical dashed line is the `Landauer-like' temperature threshold $\beta_\star$ and the diagonal dashed curve corresponds to the noise threshold $\epsilon_\star (\beta)$ for $\beta \leq \beta_\star$. The $\beta =0 $ line corresponds to the regime in which the magic protocol gives a unital channel.
    }
    \label{fig:rate_contour}
\end{figure}
\subsection{Variation of bounds with physical parameters}
The bounds derived are based on analysis of only part of the Lorenz curves and can be improved via a finer analysis.
This is apparent by simple numerical calculations on the entirety of the curves as seen in~\cref{fig:distill_bounds} for the case of unital operations.
Specifically, the existing bounds follow by considering the dominant terms in the rescaled Wigner distribution and do not, for example, take into account the Lorenz curve's peak structure.

>>>>

The above analysis makes simplifying assumptions that could easily be dropped, at the price of more complex expressions. For example, we can obtain general expressions without any assumptions on $\epsilon,\epsilon'$ or drop the Clifford change in basis. 
We could also perform similar analysis for general qudits, and different choices of magic states. It might also be of interest to consider other choices of reference states that may include more appropriate hardware physics, for example for photonic set-ups.

The only non-trivial assumption is the assumption that we can ignore correlations in the thermodynamic limit. This is a common situation in thermodynamics, where correlations are often sub-linear in the particle number in the macroscopic limit. One could perturb about this scenario, and make use of variational tools such as the Bogoliubov inequality~\cite{bogolyubov_1966} for approximating the free energy of a system, to obtain bounds in a similar form.

\begin{figure}[t!]
    \centering
    \includegraphics[scale=0.4]{figs/rate_scatter.pdf}
    \caption{\textbf{Temperature-dependent bounds for magic distillation.}
  Shown is a contour plot of  $R(\epsilon, \beta)$ for $H=H'$, where $\beta$ is an inverse temperature and $\epsilon$ is the depolarising noise of the magic state. The relevant data from the spectrum of the qutrit operator $H$ enters in via $\epsilon_\star(\beta)$ and $\beta_\star$.  The vertical dashed line is the `Landauer-like' temperature threshold $\beta_\star$ and the diagonal dashed curve corresponds to the noise threshold $\epsilon_\star (\beta)$ for $\beta \leq \beta_\star$. The $\beta =0 $ line corresponds to the regime in which the magic protocol gives a unital channel.
    }
    \label{fig:rate_contour}
\end{figure}
\subsection{Variation of bounds with physical parameters}
The bounds derived are based on analysis of only part of the Lorenz curves and can be improved via a finer analysis.
This is apparent by simple numerical calculations on the entirety of the curves as seen in~\cref{fig:distill_bounds} for the case of unital operations.
Specifically, the existing bounds follow by considering the dominant terms in the rescaled Wigner distribution and do not, for example, take into account the Lorenz curve's peak structure.
